---
title: "Rmd Template"
author: "Mike Janssen"
date: "2/23/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

In this assignment, you'll replicate one of the analyses from the week's explorations in R. You will **not** answer all the questions from the explorations, but instead synthesize those answers into the responses to the six step process outlined in the book (as well as Stat 131/132). Where appropriate, load the external datasets into this R template and use R commands to perform the required analysis. Recall the following resources for R:

* [This PDF file](https://prof.mkjanssen.org/asme/RcmdsFromClass.pdf) from Calvin University
* [The appendix to our course notes](https://prof.mkjanssen.org/asme/notes/appendix-1.html)
* [A basic guide to Markdown](https://www.markdownguide.org/basic-syntax/)


# Exploration X.Y

## Step 1: Ask a research question


## Step 2: Design a study and collect data


```{r}

```


## Step 3: Explore the data


## Step 4: Draw inferences beyond the data


## Step 5: Formulate conclusions


## Step 6: Look back and ahead



